---
title: "lab_02"
author: "derek willis"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **A1. 26 Departments. Atmospheric and Oceanic Science**

```{r}
umd_courses |>
  filter(str_detect(title, "Climate") & seats > 0) |>
  group_by(department) |>
  summarize(count_of_Classes = n()) |>
  arrange(desc(count_of_Classes))

```

**Q2.** Of the departments listed in A2, which one is the most interesting to you, and why? **A2. I think the most interesting is Geographical Science because I was really good at Geography in Elementary School.**

**Q3** Which departments have the most courses in the Fall 2023 term? And which departments in that term have at least one syllabus posted? How are the top five departments different in each result?

**A3. Business and Management with 164 courses and has the most syllabus with 82. English Department, School of Music, Biology, and Teaching don't post their syllabus on ELMS.**

```{r}
  course_count <- umd_courses |>
    filter(term == 202308) |>
    group_by(department) |>
    summarize(course =n()) |>
    arrange (desc(course))
  
  course_count_syllabus <- umd_courses |>
    filter(term == 202308 & syllabus_count >0) |>
    group_by(department) |>
    summarize(course =n()) |>
    arrange (desc(course))
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as the same dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time. **A4. Wes Moore's spending increased from 6,000 dollars to about 438,444 dollars from May 2021 to February 2022.**

```{r}
 maryland_expenses |>
  mutate(expenditure_date_new = mdy(expenditure_date)) |>
  mutate(month = floor_date(expenditure_date_new, "month")) |>
  filter(committee_name == "Moore  Wes For Maryland") |>
  group_by(month) |>
  summarize(total_amount = sum(amount))
```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for the others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **A5: I did maryland expenses, mutate in state + out of state, group bt spending location, and summarize total amount equals sum amount comma r n=T. I thought this would work because I was telling r to add in state and out of state together then to group by spending location and then summarize the total amount. I also looked at the pre-lab as well. Obviously, my equation didn't work, but this is what I thought made the most sense.**

```{r}
maryland_expenses |>
  mutate(In_state + out_of_state)|>
  group_by(spending_location) |>
  summarize(total_amount = sum(amount), na.rn=T)) |>
```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **A6. How many classes in the chemistry department have 10 seats or more. Answer = 0**

```{r}
chemistry_seats_10 <- umd_courses |> filter(department == "CHEMISTRY") |> filter(seats >= 10)

nrow(chemistry_seats_10)
```
