---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(janitor)
library(lubridate)
library(readr) 
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
donation_data <- read_csv ("data/sports-political-donations.csv")
|> clean_names() 

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: [Paste in sentence to engineer here]

-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
#this deep-pocketed group has sent $34.2 million (72.9 percent) to Republican campaigns or super PACs purely supporting Republican causes, compared to $10.1 million (21.5 percent) to Democrats over the past three elections. 
  party_donations <- donation_data |>
  mutate (
    new_party = case_when(
      Party == 'Republican' ~ 'Republican',
      Party == 'Democrat' ~ 'Democrat' ,
      .default  = 'Bipartisan'
      )
    ) |>
    mutate(new_amount = parse_number(Amount))
  
  party_donations |>
    group_by(Party) |>
    summarise(
      total_amount = sum(new_amount)
    ) |>
      arrange(desc(total_amount)) |>
      mutate(
        total_donations = sum(total_amount),
        pct_total = total_amount/total_donations
      )


# Display results of code below this codeblock

#About 72.9 percent of deep pocket dontations went to Republicans, and about 21.5 percent went to Deomcrats.A total of 34246016 donated to Republicans and a total of 10113639 donated to Democrats. We all had the same data. 





```

### Sentence 2

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
#But while that overall summary is interesting, a more pronounced signal begins to emerge when we look more closely at this group’s larger political donations. There are 40 owners who have donated at least $100,000 to Republican causes over the six-year period. 

 Republican_party_donations <- party_donations |>
   filter(new_party == "Republican") |>
    group_by(Owner) |>
    summarise(total_amount = sum(new_amount)) |>
    filter(total_amount >= 100000)
  
  


We were not able to replicate this finding because our anaylsis gave us 37 and the story says 40. We're not sure if it's an error on our part or if there's an error on there's, but it's a close number between our data and there's and it's possible there was a miscommunication.

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
#23 owners topping $100,000 to Democrats, helping lead to these splits
 Democrat_party_donations <- party_donations |>
   filter(new_party == "Democrat") |>
   group_by(Owner) |>
    summarise(total_amount = sum(new_amount)) |>
    filter(total_amount >= 100000)
  
According to the data, there were 23 Democrat Donations and on our end we got 22. Similar to the Republican data we had close numbers with the story, but we were one off this time. We're not sure if it's us with incorrect data or them, or if it's a miscommunication.'
# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
